---
title: "Data Analysis - Conflict Project"
author: "David Leydet"
date: "2022-10-22"
output:
  word_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    theme: yeti
---

This project is an exploration of climate change and conflict in Pakistan. This project uses the following datasets:

1. Pakistan Conflict Data - Acquired from the Armed Conflict Location and Event Data Project (ACLED) on 21SEP2022. See https://acleddata.com/terms-of-use/

2. Regional Climate Data - ERA5 Dataset with monthly averaged temperature and precipitation data for January and July beginning in 2010 and ending in 2022. The spatial resolution is approximately 9 kilometers. Temperature data is calculated at 2m above the land surface. Precipitation data is total precipitation between forecast steps (monthly). Reanalysis model.  See https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-land-monthly-means?tab=overview  

3. Log Gross National Income Per Capital Variable - From: https://globaldatalab.org/shdi/metadata/lgnic/ 

4. Population Data acquired on 07NOV2022 from: https://datacommons.org/place/country/PAK?utm_medium=explore&mprop=count&popt=Person&hl=en

5. Pakistan Gross Domestic Product acquired on 07NOV2022  from: https://data.worldbank.org/indicator/NY.GDP.MKTP.CD?end=2021&locations=PK&start=2010


# **Initial Data Setup**

```{r Initial Setup - Working Directory and Library, message=FALSE}

## Set working directory

setwd("~/Desktop/University of Utah PhD /Research/r_code")

## Load packages we are using

library(mapview) #mapping package
library(raster) #raster data manipulation (Climate Data)
library(RColorBrewer) #color palettes for visualization
library(sf) #simple features for spatial data
library(tmap) #mapping package
library(viridis) #color palette for visualization
library(ncdf4) #working with netCDF files (Climate Data)
library(leaflet) #basemaps for mapview
library(ggplot2) #better figures
library(ggcorrplot) #Load the correlation plot package
library(plotly) #interactive figures
library(maps) #mapping 
library(kableExtra) #creating better tables and outputs
library(dplyr) #count and data functions
library(reshape2) ## Package used to reformat data - wide to long
library(tidyverse) ##Formatting dataframes, merge, and join
library(stargazer) ##Formatting model outputs to tables

```


## **Pakistan Conflict Data**
```{r Initial Data Read - Conflict Data}

## Initial

## Pakistan Conflict Data
pak.con = read.csv("../data/conflict/pak_conflict.csv")

## Change data to a factor

#year don't set as a factor
## pak.con$year = factor(pak.con$year)

#interaction - this is the interaction code between actors (see ACLED codebook)
pak.con$interaction = factor(pak.con$interaction)

#inter1 - this is the type of actor for actor 1 (see ACLED codebook)
pak.con$inter1 = factor(pak.con$inter1)

#inter2 - this is the type of actor for actor 2(see ACLED codebook)
pak.con$inter2 = factor(pak.con$inter2)


```

```{r Convert to Spatial Data - Conflict Data, message = FALSE}

## Read in country shape file

countries = st_read("../data/conflict/ne_50m_admin_0_countries/ne_50m_admin_0_countries.shp")

## Read in administrative boundary shape file

pak.bound = st_read("../data/conflict/pak_adm_wfp_20220909_shp/pak_admbnda_adm1_wfp_20220909.shp")

pak.bound.adm2 = st_read("../data/conflict/pak_adm_wfp_20220909_shp/pak_admbnda_adm2_wfp_20220909.shp")

## Set as a spatial feature for visualization
pak.con.sf = st_as_sf(pak.con,
                      coords = c("longitude", "latitude"),
                      crs = 4326) #WGS84

```


```{r Water World Filter using DPLR}

## Using grepl to subset based on keywords
## keywords: water, rain, snow, river, flood
## ArcPro's output for the definition query is ~4400. ***What is the delta??***


pak.con.water.filter = dplyr::filter(pak.con.sf, grepl("water|flood|rain|river|snow", pak.con$notes, ignore.case = TRUE))

str(pak.con.water.filter)

```

```{r Save Water Conflict Shapefile, warning = FALSE, eval = FALSE}

pak.con.water.filter.sf = st_as_sf(pak.con.water.filter)
st_write(pak.con.water.filter.sf,
         dsn = "./figures/pak_water_con_sf.shp",
         delete_layer = TRUE) ##Overrite the existing file


```


## **Initial Visualization Exploration**

```{r Initial Visualization - Conflict Data}

## Total events
ggplot(pak.con.sf)  +
  geom_sf(color = "firebrick3", alpha = 0.4)


```

```{r Initial Visualization - tmap}

con.total.map = 
  tm_shape(countries , bbox = st_bbox(pak.con.sf)) +
  tm_borders(col = "gray") +
  tm_shape(pak.con.sf) +
  tm_symbols(col = "event_type", 
             alpha = 0.4,
             size = 0.2,
             title.col = "Conflicts By Event Type") +
   tm_layout(main.title = "Pakistan Conflicts: 2010 - 2022",
            legend.outside = TRUE)

con.total.map
```

```{r Map Save - Total Conflicts, message = FALSE, eval=FALSE}

pdf("./figures/pak_conflict_total")
con.total.map
dev.off()

```

```{r Water Conflicts}

con.water.map = 
  tm_shape(countries , bbox = st_bbox(pak.con.sf)) +
  tm_borders(col = "gray") +
  tm_shape(pak.con.water.filter) +
  tm_symbols(col = "event_type", 
             alpha = 0.4,
             size = 0.2,
             title.col = "Water Conflicts By Event Type") +
   tm_layout(main.title = "Pakistan Water Conflicts: 2010 - 2022",
            legend.outside = TRUE)

con.water.map

```

```{r By Year Look - Water Conflicts}

water.con.by.year.map = 
  tm_shape(countries , bbox = st_bbox(pak.con.sf)) +
  tm_borders(col = "gray") +
  tm_shape(pak.con.water.filter) +
  tm_facets(by = "year") +
  tm_symbols(col = "dodgerblue1", 
             alpha = 0.5,
             size = 0.2) +
   tm_layout(main.title = "Pakistan Water Conflicts: 2010 - 2022",
             main.title.size = 0.75,
             legend.outside = FALSE,
             legend.position = c("right", "bottom"))

water.con.by.year.map


```

```{r Map Save - Water Conflict by Year, eval=FALSE, message=FALSE}

pdf("./figures/pak_water_con_by_year")
water.con.by.year.map
dev.off()

```

```{r By Year Look - Conflict}

con.by.year.map = 
  tm_shape(countries , bbox = st_bbox(pak.con.sf)) +
  tm_borders(col = "gray") +
  tm_shape(pak.con.sf) +
  tm_facets(by = "year") +
  tm_symbols(col = "firebrick3", 
             alpha = 0.5,
             size = 0.2) +
   tm_layout(main.title = "Pakistan Conflicts: 2010 - 2022",
            legend.outside = TRUE)

con.by.year.map

```

```{r Map Save - Conflict by Year, eval=FALSE, message=FALSE}

pdf("./figures/pak_con_by_year")
con.by.year.map
dev.off()

```

```{r Visualize by fatlities - Conflict}

# Subset to exclude 0 fatality events

pak.con.sf.fat = subset(pak.con.sf, fatalities > 10)

## NOT WORKING PROPERLY*****
tm_shape(countries , bbox = st_bbox(pak.con.sf.fat)) +
  tm_borders(col = "gray") +
  tm_shape(pak.con.sf.fat) +
  tm_symbols(col = "fatalities",
             palette = "Reds",
             alpha = 0.4,
             title.col = "Number of Fatalities",
             size = 0.2) +
   tm_layout(main.title = "Pakistan Conflicts: 2010 - 2022",
            legend.outside = TRUE) 
   

```


```{r Pakistan Admin Level 1 Boundary Shapefile}

tm_shape(pak.bound) +
  tm_borders() +
  tm_fill(col = "ADM1_EN",
          title = "Pakistan Provinces")


```

```{r Mapping Conflict Across Pakistan and its Provinces}

## Generate a new color palette
mycol = turbo(n = 8, begin = 0.5, end = 1, direction = 1)

con.by.district = 
  tm_shape(pak.bound) +
  tm_borders() +
  tm_fill(col = "ADM1_EN",
          title = "Pakistan Provinces") +
  tm_shape(pak.con.sf.fat) +
  tm_symbols(col = "fatalities",
             palette = mycol,
             alpha = 0.4,
             title.col = "Number of Fatalities",
             size = 0.2) +
   tm_layout(main.title = "Pakistan Conflicts by Province: 2010 - 2022",
            legend.outside = TRUE) +
   tm_compass(position = c("left", "top"))

con.by.district
```


```{r Map Save - Conflict by District, message = FALSE, eval=FALSE}

pdf("./figures/con_by_district")
con.by.district
dev.off()

```


```{r Pakistan Admin Level 2 Boundary Shapefile, warning=FALSE}

tm_shape(pak.bound.adm2) +
  tm_borders() +
  tm_fill(col = "ADM2_EN",
          title = "Pakistan Divisions") +
  tm_layout(legend.show = FALSE)


```



## **Initial Data Analysis**

```{r Number of Conflicts by Year, warning=FALSE}

## Create the counts by year

year.2022 = as.numeric(nrow(filter(pak.con, year == 2022)))
year.2021 = as.numeric(nrow(filter(pak.con, year == 2021)))
year.2020 = as.numeric(nrow(filter(pak.con, year == 2020)))
year.2019 = as.numeric(nrow(filter(pak.con, year == 2019)))
year.2018 = as.numeric(nrow(filter(pak.con, year == 2018)))
year.2017 = as.numeric(nrow(filter(pak.con, year == 2017)))
year.2016 = as.numeric(nrow(filter(pak.con, year == 2016)))
year.2015 = as.numeric(nrow(filter(pak.con, year == 2015)))
year.2014 = as.numeric(nrow(filter(pak.con, year == 2014)))
year.2013 = as.numeric(nrow(filter(pak.con, year == 2013)))
year.2012 = as.numeric(nrow(filter(pak.con, year == 2012)))
year.2011 = as.numeric(nrow(filter(pak.con, year == 2011)))
year.2010 = as.numeric(nrow(filter(pak.con, year == 2010)))

## Create a new data frame for the counts

# vector of years

years = c("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021", "2022")

# vector of years as a number for plotting purposes
years.num = seq(from = 2010, to = 2022, by = 1)

# vector of conflict counts

num.conflicts = c(year.2010, year.2011, year.2012, year.2013, year.2014, year.2015, year.2016, year.2017, year.2018, year.2019, year.2020, year.2021, year.2022)

# Combine the vectors into a data frame

con.by.year = data.frame (years, num.conflicts)

# Combine the vectors using the years as a number

con.by.year.num = data.frame(years.num, num.conflicts)

# Visualize the Count
# How does one add a trendline?

num.conflicts.sctrplot =
  ggplot(data = con.by.year.num, aes(x = years.num, y = num.conflicts)) +
  geom_point(shape = 18, size = 3, color = "red") +
  geom_smooth(method = lm,
              se = FALSE) +
  xlab("Year") +
  ylab("Number of Conflicts") +
  labs(title = "Pakistan Conflicts by Year") +
  scale_x_discrete(limits = c(years.num)) +
  theme_bw() 

num.conflicts.sctrplot
```

```{r Scatter Plot Save - Conflict by Year, message = FALSE, eval=FALSE}

pdf("./figures/con_by_year_scatterplot")
num.conflicts.sctrplot
dev.off()

```


```{r Number of Water Conflicts by Year}

##Count the number of water conflicts by year
water.year.2022 = as.numeric(nrow(filter(pak.con.water.filter, year == 2022)))
water.year.2021 = as.numeric(nrow(filter(pak.con.water.filter, year == 2021)))
water.year.2020 = as.numeric(nrow(filter(pak.con.water.filter, year == 2020)))
water.year.2019 = as.numeric(nrow(filter(pak.con.water.filter, year == 2019)))
water.year.2018 = as.numeric(nrow(filter(pak.con.water.filter, year == 2018)))
water.year.2017 = as.numeric(nrow(filter(pak.con.water.filter, year == 2017)))
water.year.2016 = as.numeric(nrow(filter(pak.con.water.filter, year == 2016)))
water.year.2015 = as.numeric(nrow(filter(pak.con.water.filter, year == 2015)))
water.year.2014 = as.numeric(nrow(filter(pak.con.water.filter, year == 2014)))
water.year.2013 = as.numeric(nrow(filter(pak.con.water.filter, year == 2013)))
water.year.2012 = as.numeric(nrow(filter(pak.con.water.filter, year == 2012)))
water.year.2011 = as.numeric(nrow(filter(pak.con.water.filter, year == 2011)))
water.year.2010 = as.numeric(nrow(filter(pak.con.water.filter, year == 2010)))

##Automate this: function script?
##Function Works!
##Automate loop??

mycountbyyear = function(x) {
  as.numeric(nrow(filter(x, year == 2010)))
}

testyear = mycountbyyear(pak.con.water.filter)


##Create a vector of water conflicts by year
water.num.conflicts = c(water.year.2010, water.year.2011, water.year.2012, water.year.2013, water.year.2014, water.year.2015, water.year.2016, water.year.2017, water.year.2018, water.year.2019, water.year.2020, water.year.2021, water.year.2022)

# Combine the vectors into a data frame. Need to use years.num in order to plot a trend line

water.con.by.year = data.frame (years.num, water.num.conflicts)

##Plot it!

num.water.conflicts.sctrplot =
  ggplot(data = water.con.by.year, aes(x = years.num, y = water.num.conflicts)) +
  geom_point(shape = 18, size = 3, color = "blue") +
  geom_smooth(method = NULL,
              se = TRUE,
              color = "black",
              linetype = "dashed",
              size = 0.5) +
  xlab("Year") +
  ylab("Number of Water Conflicts") +
  labs(title = "Pakistan Water Conflicts by Year") +
  scale_x_discrete(limits = c(years.num)) +
  theme_bw() 

num.water.conflicts.sctrplot


```


```{r Scatter Plot Save - Water Conflict by Year, message = FALSE, eval=FALSE}

pdf("./figures/water_con_by_year_scatterplot")
num.water.conflicts.sctrplot
dev.off()

```


## **Pakistan GNI Data**

```{r Initial Data Read - GNI}

## Initial Read
## ****Note - the years are from 1990 - 2019****
## Log Gross National Income per capita in thousands of US Dollars (2011 PPP) 

## Ask Simon about time series data

pak.gni = read.csv("../data/conflict/pak_subnat_gdp.csv")

## Convert it from wide format to long format
pak.gni.long = melt(pak.gni, id.vars = "Region")

## Change the column name to year
colnames(pak.gni.long)[2] = "year" 

## Remove the "x" from the years
pak.gni.long$year = gsub("X", "", as.factor(pak.gni.long$year))

## Subset the temporal timeframe: 2010 - 2022 (2019 in this case)

###pak.gni.time.subset = subset(pak.gni.long, Factor == "2010")



```


```{r GNI Visualization}

pak.gni.plot = 
  ggplot(data = pak.gni.long, aes(x = year, y = value, color = Region)) +
  geom_point() +
  xlab("Year") +
  ylab("GNI per Capita (Thousands USD)") +
  # xlim(2010, 2019) +
  theme(axis.text.x = element_text(angle = 90))
  theme_bw()
  

pak.gni.plot


```

```{r GNI Timeseries Plot Save, message = FALSE}

## Save as a .pdf
ggsave("./figures/gni_by_year_district.pdf", pak.gni.plot)


```


```{r Initial Pakistan GDP Read}

##2022 GDP has not been published yet, so we estimated based on an average gdp growth of 3.7%

pak.gdp = read.csv("../data/conflict/pak_gdp_2010_2022.csv")

pak.gdp.plot = 
  ggplot(data = pak.gdp, aes(x = year, y = gdp)) +
  geom_point(col = "forestgreen") +
  geom_line(linetype = "dashed",
            col = "forestgreen") +
  labs(title = "Pakistan GDP (2010-2022)") +
  xlab("Year") +
  ylab("GDP USD($)")

pak.gdp.plot
#May want to scale the GDP

```

```{r Scatter Plot Save - Pakistan GDP, message = FALSE, eval=FALSE}

pdf("./figures/pak_gdp_by_year")
pak.gdp.plot
dev.off()

```

## **Pakistan Climate Data**

### **Temperature Data**

```{r Initial Data Read - Climate: Temperature }

## Read in the ERA5 climate data

pak.temp = raster("../data/conflict/pak_clim.nc", varname = "t2m")

## Look at the data
##26 bands (layers)
##13 years worth of data * 2 months(Jan, Jul) = 26 observations (bands?)

## Syntax variable = raster("filepath", band = #) to extract the specific band you need

## library(ncdf4) this package allows you to read into the metadata
## Could use stack() to look at all of the bands

pak.temp

## CRS Check
## Coordinate Reference System:
## Deprecated Proj.4 representation: +proj=longlat +datum=WGS84 +no_defs 


```


```{r Initial Temperature Visualization}

## Temperature is in Kelvin (how do I convert the scale to Celsius? Kelvin Temp - 273.15)
## January Temperature. How to plot July?

mytemp.pal = brewer.pal(n = 9, name = "OrRd")

plot(pak.temp, 
     main = "January Temperature - 2010",
     col = mytemp.pal)
plot(st_geometry(pak.bound),
     border = "black",
     add = TRUE)



```



```{r Masking Visualization - Test}

##Not working correctly now? 20221107
pak.temp.masked = mask(pak.temp, mask = pak.bound)

plot(pak.temp.masked,
     main = "Pakistan Jan 2010 Temperature - Masked",
     col = mytemp.pal)
plot(st_geometry(pak.bound), add = TRUE)


```

```{r Temperature Data Extraction - Stack}

##Use stack() to load all of the temperature bands

temp.stack = stack("../data/conflict/pak_clim.nc", varname = "t2m")

##View the stack data
temp.stack

```


```{r Temperature Data Stack Subset}

##January 2010 stack
temp.stack.jan2010 = temp.stack[[1]]

temp.stack.jan2010

```

```{r Crop the Stack by Pakistan}

##Crop the temperature stack by the Pakistan Boundary

temp.stack.crop = crop(temp.stack, pak.bound)

temp.stack.crop

```


```{r Temperature Stack Visualization}
##Color Template for Temperature


##Create a function to add the Pakistan Level 2 boundaries to the figure
addborder = function(){
  plot(as_Spatial(pak.bound), 
       add = TRUE)
}

plot(temp.stack.crop[[1]],
     col = mytemp.pal,
     zlim = c(250, 320), ##zlim makes sure the scale is the same
     addfun = addborder)


```



```{r Mask the Stack for Pakistan}

pak.temp.stack.mask = mask(temp.stack, mask = pak.bound)

plot(pak.temp.stack.mask,
     col = mytemp.pal,
     zlim = c(250, 320), ##zlim makes sure the scale is the same
     addfun = addborder)

```


```{r Temperature Stack Mean Temperature}

##Use the temperature stack (masked) to calculate the mean average temperature for the country

temp.avg.all = cellStats(pak.temp.stack.mask, mean)

##Visualize the data - this differs slighlt from running the entire raster data (without the mask)
#plot(temp.avg.all,
     #type = "p")

temp.avg.all

##Shape the average temperature values into a data frame
temp.avg.df = as.data.frame(temp.avg.all)

temp.avg.df

```


```{r Write to CSV - Temperature Average, eval=FALSE}

##Write to csv to transform....need to find a better way to do this!!!!***
write.csv(temp.avg.df,
          "../data/conflict/pak_avg_temp")


```


```{r Temperature Read Converted csv}
##Read in the converted csv

pak.temp.avg = read.csv("../data/conflict/pak_avg_temp_converted.csv")

pak.temp.avg

```

```{r Temperature Trend Visualization}

avg.temp.plot =
  ggplot(data = pak.temp.avg) +
  geom_line(aes(x=year, y = avg_jan_celsius), col = "blue") +
   geom_line(aes(x=year, y = avg_jul_celsius), col = "red") +
  xlab("Year") +
  ylab("Temperature (Celsius)")


avg.temp.plot

```


### **Precipitation Data**


```{r Initial Data Read - Climate: Precipitation }

## Read in the ERA5 climate data
## Value is in meters

##26 bands (layers)
##13 years worth of data * 2 months(Jan and Jul) = 26 observations (bands?)

## Syntax variable = raster("filepath", band = #) to extract the specific band you need

## library(ncdf4) this package allows you to read into the metadata
## Could use stack() to look at all of the bands

pak.precip = raster("../data/conflict/pak_clim.nc", varname = "tp")

## Look at the data

pak.precip


```


```{r Initial Precipitation Visualization}

##January precipitation
myprecip.pal = brewer.pal(n = 9, name = "Blues")
plot(pak.precip, 
     main = "January Precipitation - 2010",
     col = myprecip.pal)
plot(st_geometry(pak.bound.adm2), 
     border = "black",
     add = TRUE)


```


```{r Precipitation Data Extraction - Stack}

##Use stack() to load all of the precipitation bands
##REMINDER This is total precipitation in meters


precip.stack = stack("../data/conflict/pak_clim.nc", varname = "tp")

##View the stack data
precip.stack

```

```{r Precip Mask the Stack for Pakistan}

pak.precip.stack.mask = mask(precip.stack, mask = pak.bound)

plot(pak.precip.stack.mask,
     col = myprecip.pal,
     zlim = c(0, 0.0030), ##zlim makes sure the scale is the same
     addfun = addborder)

```


```{r Precip Stack Mean}

##Use the precipitation stack (masked) to calculate the mean average temperature for the country

precip.avg.all = cellStats(pak.precip.stack.mask, mean)

##Visualize the data - this differs slighlt from running the entire raster data (without the mask)
#plot(precip.avg.all,
     #type = "p")

precip.avg.all

##Shape the average temperature values into a data frame
precip.avg.df = as.data.frame(precip.avg.all)

precip.avg.df

```


```{r Write to CSV - Precipitation Average, eval=FALSE}

##Write to csv to transform....need to find a better way to do this!!!!***
write.csv(precip.avg.df,
          "../data/conflict/pak_avg_precip")


```


```{r Precipitation Read Converted csv}
##Read in the converted csv

pak.precip.avg = read.csv("../data/conflict/pak_avg_precip_converted.csv")

pak.precip.avg

```


```{r Precipitation Trend Visualization}

avg.precip.plot =
  ggplot(data = pak.precip.avg) +
  geom_line(aes(x=year, y = avg_jan_precip_mm, color = "avg_jan_precip")) +
   geom_line(aes(x=year, y = avg_jul_precip_mm, color = "avg_jul_precip")) +
  xlab("Year") +
  ylab("Total Precipitation (mm)") +
  scale_color_manual(name = "Seasonal Precipitation",
                     values = c("avg_jan_precip" = "blue", 
                                "avg_jul_precip" = "red"))


avg.precip.plot

```

## **Population Data**

```{r Read Population Data}

##Data acquired from datacommons.org (world bank)
##2022 is estimated based on the 10-year average growth rate


pak.pop = read.csv("../data/conflict/pak_pop.csv")

ggplot(data = pak.pop, aes(x = year, y = pak_pop)) +
  geom_point(col = "darkorange2") +
  geom_line(linetype = "dashed",
            col = "darkorange2") +
  labs(title = "Pakistan Population (2010-2022)") +
  xlab("Year") +
  ylab("Population")


```



# **Models Build**

```{r Create a New Data Frame - National Model}

##Test: Merge the GDP and Number of Conflicts together

##Change the column header in order to merge
colnames(con.by.year)[1] = "year"

##Merge
pak.model.df = merge(con.by.year, pak.gdp, by = "year")
##It worked!

##Add water conflicts 
##use by.x and by.y to merge using different variable names
pak.model.df = merge(pak.model.df, water.con.by.year, by.x ="year", by.y = as.character("years.num"))

##Now to add population
pak.model.df = merge(pak.model.df, pak.pop, by = "year")


##Convert year to number

pak.model.df$year = as.numeric(pak.model.df$year)

##Merge temperature data

pak.model.df = merge(pak.model.df, pak.temp.avg, by = "year")

##Merge precipitation data

pak.model.df = merge(pak.model.df, pak.precip.avg, by = "year")



```


```{r Write to CSV - Conflict Model Dataframe, eval=FALSE}

##Write to csv to save***
write.csv(pak.model.df,
          "../data/conflict/pak_model_df")


```


```{r Correlation Matrix - National}

#pairs(pak.model.df)

ggcorrplot(cor(pak.model.df), 
           method = "square",
           type = "lower",
           lab = TRUE,
           colors = c("blue", "darksalmon", "firebrick"))
           #p.mat = cor_pmat(pak.model.df)) #x out non-significant p-values

```


## **OLS LM**

```{r OLS LM Build}

pak.nat.lm = lm(num.conflicts ~ year + gdp + pak_pop + avg_jan_celsius + avg_jul_celsius + avg_jan_precip_mm + avg_jul_precip_mm,
                data = pak.model.df)

summary(pak.nat.lm)

```


## **GLM - Poisson**

```{r Model Build - National}

##Starting with a generalized linear model - Poisson - as the number of conflicts is count data
## Try running it without year
## Running the Z-score for the variables (the interpretation would be the change in one standard deviation) - use function scale()

pak.nat.glm = glm(num.conflicts ~ year + gdp + pak_pop + avg_jan_celsius + avg_jul_celsius + avg_jan_precip_mm + avg_jul_precip_mm,
              family = poisson(link = 'log'),
              data = pak.model.df)

summary(pak.nat.glm)

##Need to transform the coefficients 
```

```{r Stargazer Conversion Model Output Table}
##Create a variable to store the model summary

#pak.nat.glm.sum = summary(pak.nat.glm)

#stargazer(pak.nat.glm.sum, type = "html", style = "all")

```


```{r Transform the Coefficients}

##Coefficient transformation as the output is in the log values (link = log)
##Reminder Interpretation Note:
##Log odds scale to odds scale - it becomes multiplicative - a rate of change (bigger than 1 it is a positive change, if it is less than 1 it is a negative change)
##Scale the gdp coefficient and population

pak.nat.glm.coef = exp(coef(pak.nat.glm))
pak.nat.glm.coef


```

### **Residuals**

```{r Residual Visualization}

##View the residuals

plot(residuals.glm(pak.nat.glm))
abline(h = 0, col = "red")

##Fitting a spline may be better for this model?
```


```{r Model Plot}

plot(pak.nat.glm)


```








