---
title: "HMA Data Analysis (Clean)"
author: "David Leydet"
date: "2024-04-18"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    theme: yeti
---

# **Introduction**

This document consolidates the modeling and analysis efforts into one document. The data used in this analysis is cleaned prior to importing.


```{r Set Working Directory, echo=FALSE}

setwd("~/Desktop/University of Utah PhD /Research/r_code/")

```


# **Administrative Level 1 Analysis (Imputed Data)**

## **Imputed Data Read and Adjustment**

```{r Adm1 Data}
# load cleaned data
adm.imp = read.csv("../cleaned_data/hma_adm1_imp_df_clean_20240418.csv")

# check
str(adm.imp)

# remove row number variable (X) from df
adm.imp = adm.imp %>% 
  select(-X)

# factor country and adm 1
adm.imp$country = as.factor(adm.imp$country)
adm.imp$adm1 = as.factor(adm.imp$adm1)

# check
str(adm.imp)
```

```{r Adm1 Zscore}
# z score the conflict variable
adm.imp = adm.imp %>% 
  mutate(zscore = scale(con, center = TRUE, scale = TRUE))

# Check
str(adm.imp)

```

## **Initial Data Visualizations**

```{r Adm 1 visual}
plot(adm.imp$year, adm.imp$con)

```


```{r Adm 1 Histogram of variables}
## Check the histogram of variables for normality

## Gross domestic product proxy - nighttime lights
h1 = adm.imp %>% 
  gghistogram(x = "nlights")

## Population (AIDDATA)
h2 = adm.imp %>% 
  gghistogram(x = "pop")

## Temp (AIDDATA)
h3 = adm.imp %>% 
  gghistogram(x = "temp")

## Precip (AIDDATA)
h4 = adm.imp %>% 
  gghistogram(x = "precip")

## Total Conflict count (ACLED)
h5 = adm.imp %>% 
  gghistogram(x = "con")

## Visualize 
ggarrange(h1, h2, h3, h4, h5)

```

```{r Adm 1 Histogram of variables 2}
## Check the histogram of variables for normality
# World Governance Indicators

h6 = adm.imp %>% 
  gghistogram(x = "ge")


h7 = adm.imp %>% 
  gghistogram(x = "cpt")


h8 = adm.imp %>% 
  gghistogram(x = "rol")


h9 = adm.imp %>% 
  gghistogram(x = "rq")


h10 = adm.imp %>% 
  gghistogram(x = "polstab")

h11 = adm.imp %>% 
  gghistogram(x = "voice")

## Visualize 
ggarrange(h6, h7, h8, h9, h10, h11)

```


## **Random Forest**

```{r Adm1 RF iml}
# load libraries
library(randomForest) #rf package
library(iml) #interpretable machine learning package

```

```{r Adm1 RF Data Frame}
# subset the variables we want to use for this model
# remove zscore and adm1 (adm1 has too many factors)
adm.imp.rf.df = adm.imp %>% 
  select(-zscore, -adm1)

str(adm.imp.rf.df)

```

```{r Adm 1 RF Train}
st = Sys.time()

# Train the random forest 
set.seed(123)
rf = randomForest(con ~ .,
                  data = adm.imp.rf.df)

et = Sys.time()
elapsed = et-st
print(elapsed)
```

```{r Adm1 RF Errors}
##pseudo r2 value
mean(rf$rsq)

```

### *iml Objects and Interpretations**

```{r Adm1 Predictor Object}
## Create a data frame with the features minus the target
x = adm.imp.rf.df %>% 
  select(-con)

# Store the data and con in the predictor container along with the randomForest parameters
adm.imp.predictor = Predictor$new(model = rf,
                                  data = x,
                                  y = adm.imp.rf.df$con)

```


```{r Adm1 Feature Importance, message=FALSE}
## Store the features in a FeatureImp object
## loss argument specifies the performance measure for error
importance = FeatureImp$new(predictor = adm.imp.predictor,
                     loss = "rmse")

```


```{r Adm1 Importance Visualization}
## Visualize the importance using ggplot2
library(ggplot2)

##Plot
plot(importance)

```

**Feature Importance**

```{r Adm1 Importance Results}
## View the the feature importance percentiles
importance$results

```


**Feature Effects**

```{r Adm1 ALE Plot}
## Reminder this allows us to interpret areas of the curve that may be more important than others
##grid.size argument is the number of quantiles specified

ale = FeatureEffect$new(predictor = adm.imp.predictor,
                        feature = "pop",
                        grid.size = 10)

ale$plot()
```

```{r Adm1 ALE Plot Loop}
## Define a helper function to plot

ale.plot = function(x){
  ale$set.feature(x)
  ale$plot()
}

```

```{r Adm1 ALE Plots Complete}
# plot through the variables
ale.plot("country")
ale.plot("temp")
ale.plot("polstab")
ale.plot("precip")
ale.plot("voice")
ale.plot("ge")
ale.plot("cpt")
ale.plot("nlights")
ale.plot("rol")
ale.plot("rq")
ale.plot("year")
ale.plot("pop")

```

```{r Adm 1 Plot Feature Effects - All}
##Plot all of the feature effects at once

effs = FeatureEffects$new(predictor = adm.imp.predictor,
                          grid.size = 10)

# Visualize
plot(effs)

```

**Partial Dependecy Plots**

**Note:** "Individual Conditional Expectation (ICE) plots display one line per instance that shows how the instance’s prediction changes when a feature changes.

The partial dependence plot for the average effect of a feature is a global method because it does not focus on specific instances, but on an overall average. The equivalent to a PDP for individual data instances is called individual conditional expectation (ICE) plot (Goldstein et al. 201748)."

The yellow line represents the partial dependency of the feature, which is the average of the ICE plots.

```{r Adm1 PDPs, message=FALSE, warning=FALSE}
#population
#store it in an object
rf.pop = Partial$new(predictor = adm.imp.predictor,
                        feature = "pop",
                        aggregation = "pdp",
                        ice = TRUE)

rf.country = Partial$new(predictor = adm.imp.predictor,
                        feature = "country",
                        aggregation = "pdp",
                        ice = TRUE)

rf.temp = Partial$new(predictor = adm.imp.predictor,
                        feature = "temp",
                        aggregation = "pdp",
                        ice = TRUE)

rf.precip = Partial$new(predictor = adm.imp.predictor,
                        feature = "precip",
                        aggregation = "pdp",
                        ice = TRUE)

rf.nlights = Partial$new(predictor = adm.imp.predictor,
                        feature = "nlights",
                        aggregation = "pdp",
                        ice = TRUE)

rf.ge = Partial$new(predictor = adm.imp.predictor,
                        feature = "ge",
                        aggregation = "pdp",
                        ice = TRUE)

rf.cpt = Partial$new(predictor = adm.imp.predictor,
                        feature = "cpt",
                        aggregation = "pdp",
                        ice = TRUE)

rf.rol = Partial$new(predictor = adm.imp.predictor,
                        feature = "rol",
                        aggregation = "pdp",
                        ice = TRUE)

rf.rq = Partial$new(predictor = adm.imp.predictor,
                        feature = "rq",
                        aggregation = "pdp",
                        ice = TRUE)

rf.polstab = Partial$new(predictor = adm.imp.predictor,
                        feature = "polstab",
                        aggregation = "pdp",
                        ice = TRUE)

rf.voice = Partial$new(predictor = adm.imp.predictor,
                        feature = "voice",
                        aggregation = "pdp",
                        ice = TRUE)

# center (this centers the impact of y hat on a starting value)
rf.pop$center(min(adm.imp.rf.df$pop))
rf.temp$center(min(adm.imp.rf.df$temp))
rf.precip$center(min(adm.imp.rf.df$precip))
rf.nlights$center(min(adm.imp.rf.df$nlights))
rf.ge$center(min(adm.imp.rf.df$ge))
rf.cpt$center(min(adm.imp.rf.df$cpt))
rf.rol$center(min(adm.imp.rf.df$rol))
rf.rq$center(min(adm.imp.rf.df$rq))
rf.polstab$center(min(adm.imp.rf.df$polstab))
rf.voice$center(min(adm.imp.rf.df$voice))


```


```{r Adm1 PDPs 2}
# plot
p1 = plot(rf.pop) + ggtitle("Population PDP/ICE Plot")
p2 = plot(rf.country) + ggtitle("Country PDP/ICE Plot")
p3 = plot(rf.temp) + ggtitle("Temperatire PDP/ICE Plot")
p4 = plot(rf.precip) + ggtitle("Precipitation PDP/ICE Plot")

gridExtra::grid.arrange(p1, p2, p3, p4, nrow = 2)

```


```{r Adm1 PDPs 3}
# plot
p5 = plot(rf.nlights) + ggtitle("Nighttime Lights PDP/ICE Plot")
p6 = plot(rf.ge) + ggtitle("Governance Effectiveness PDP/ICE Plot")
p7 = plot(rf.cpt) + ggtitle("Corruption PDP/ICE Plot")
p8 = plot(rf.rol) + ggtitle("Rule of Law PDP/ICE Plot")

gridExtra::grid.arrange(p5, p6, p7, p8, nrow = 2)

```

```{r Adm1 PDPs 4}
# plot
p9 = plot(rf.rq) + ggtitle("Regulatory Quality PDP/ICE Plot")
p10 = plot(rf.polstab) + ggtitle("Political Stability PDP/ICE Plot")
p11 = plot(rf.voice) + ggtitle("Voice and Accountability PDP/ICE Plot")


gridExtra::grid.arrange(p9, p10, p11, nrow = 2)

```


**Measure Interactions**

```{r Adm 1 Interactions}
st = Sys.time()
##Set up the interactions wrapper
##Play around with grid.size

adm.interact = Interaction$new(predictor = adm.imp.predictor,
                           grid.size = 15)

et = Sys.time()
print(et-st)
```

```{r Adm 1 Plot Interactions}
##Plot the features to see how the interact with any other feature in the data
# scale 0 - 1 with 1 meaning that 100% of the variance is explained with interactions with the other features
plot(adm.interact)

```


## **Surrogate Model**

```{r Adm 1 Surrogate Model}

tree = TreeSurrogate$new(predictor = adm.imp.predictor,
                         maxdepth = 2)


# Visualize
plot(tree)

```


```{r Adm 1 Plot The RF Tree}

# Use this code to plot the random forest tree itself
plot(tree$tree)

```


```{r Adm 1 Surrogate Model Evaluation} 
## Check the model performance 

print(tree$r.squared)

# 0.47 - doesn't perform that well

```





# **Administrative Level 1 Analysis (Non-Imputed Data)**

This analysis re-runs the model code on the non-imputed data frame to see if there is any noticeable differences in performance or outcomes.


```{r Non-imp data read}
## read in the non-imputed data
adm.nonimp = read.csv("../cleaned_data/hma_adm1_non_imp_df_clean_20240418.csv")

# remove the "X" column
adm.nonimp = adm.nonimp %>% 
  select(-X)

# factor country and adm 1
adm.nonimp$country = as.factor(adm.nonimp$country)
adm.nonimp$adm1 = as.factor(adm.nonimp$adm1)

# check 
str(adm.nonimp)

```

## **Random Forest (Non-Imputed Data)**


```{r Non-Imp RF Data Frame}
# subset the variables we want to use for this model
# remove adm1 (adm1 has too many factors)
adm.nonimp.rf.df = adm.nonimp %>% 
  select(-adm1)

str(adm.nonimp.rf.df)

```

```{r Non-Imp RF Train}
st = Sys.time()

# Train the random forest 
set.seed(456)
nonimp.rf = randomForest(con ~ .,
                  data = adm.nonimp.rf.df)

et = Sys.time()
elapsed = et-st
print(elapsed)
```

```{r Non-Imp RF Errors}
##pseudo r2 value
mean(nonimp.rf$rsq)

```

```{r Performance Comparison}
# Imputed model performance
mod1.rsq = mean(rf$rsq)

# Non-imputed model performance
mod2.rsq = mean(nonimp.rf$rsq)

print(c(mod1.rsq, mod2.rsq))

```

### *Non-Imp iml Objects and Interpretations**

```{r Non Imp Predictor Object}
## Create a data frame with the features minus the target
y = adm.nonimp.rf.df %>% 
  select(-con)

# Store the data and con in the predictor container along with the randomForest parameters
adm.nonimp.predictor = Predictor$new(model = nonimp.rf,
                                  data = y,
                                  y = adm.nonimp.rf.df$con)

```

**Feature Importance**

```{r Non Imp Feature Importance, message=FALSE}
## Store the features in a FeatureImp object
## loss argument specifies the performance measure for error
nonimp.importance = FeatureImp$new(predictor = adm.nonimp.predictor,
                     loss = "rmse")

```


```{r Non Imp Importance Visualization}

##Plot
plot(nonimp.importance)

```

**ALE Plots**

```{r Non Imp Plot Feature Effects - All}
##Plot all of the feature effects at once

nonimp.effs = FeatureEffects$new(predictor = adm.nonimp.predictor,
                          grid.size = 10)

# Visualize
plot(nonimp.effs)

```


**Measure Interactions**

```{r Non Imp Interactions}
st = Sys.time()
##Set up the interactions wrapper
##Play around with grid.size

nonimp.interact = Interaction$new(predictor = adm.nonimp.predictor,
                           grid.size = 15)

et = Sys.time()
print(et-st)
```

```{r Non Imp Plot Interactions}
##Plot the features to see how the interact with any other feature in the data
# scale 0 - 1 with 1 meaning that 100% of the variance is explained with interactions with the other features
plot(nonimp.interact)

```








