---
title: "HMA Data Analysis (Clean)"
author: "David Leydet"
date: "2024-04-18"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    theme: yeti
---

# **Introduction**

This document consolidates the modeling and analysis efforts into one document. The data used in this analysis is cleaned prior to importing.


```{r Set Working Directory, echo=FALSE}

setwd("~/Desktop/University of Utah PhD /Research/r_code/")

```


# **Administrative Level 1 Analysis (Imputed Data)**

## **Imputed Data Read and Adjustment**

```{r Adm1 Data}
library(tidyverse)

# load cleaned data
adm.imp = read.csv("../cleaned_data/hma_adm1_imp_df_clean_20240418.csv")

# check
str(adm.imp)

# remove row number variable (X) from df
adm.imp = adm.imp %>% 
  select(-X)

# factor country and adm 1
adm.imp$country = as.factor(adm.imp$country)
adm.imp$adm1 = as.factor(adm.imp$adm1)

# check
str(adm.imp)
```

```{r Adm1 Zscore}
# z score the conflict variable
adm.imp = adm.imp %>% 
  mutate(zscore = scale(con, center = TRUE, scale = TRUE))

# Check
str(adm.imp)

```

## **Initial Data Visualizations**

```{r Adm 1 visual}
plot(adm.imp$year, adm.imp$con)

```


```{r Adm 1 Histogram of variables}
library(ggplot2)
library(ggpubr)
## Check the histogram of variables for normality

## Gross domestic product proxy - nighttime lights
h1 = adm.imp %>% 
  gghistogram(x = "nlights")

## Population (AIDDATA)
h2 = adm.imp %>% 
  gghistogram(x = "pop")

## Temp (AIDDATA)
h3 = adm.imp %>% 
  gghistogram(x = "temp")

## Precip (AIDDATA)
h4 = adm.imp %>% 
  gghistogram(x = "precip")

## Total Conflict count (ACLED)
h5 = adm.imp %>% 
  gghistogram(x = "con")

## Visualize 
ggarrange(h1, h2, h3, h4, h5)

```

```{r Adm 1 Histogram of variables 2}
## Check the histogram of variables for normality
# World Governance Indicators

h6 = adm.imp %>% 
  gghistogram(x = "ge")


h7 = adm.imp %>% 
  gghistogram(x = "cpt")


h8 = adm.imp %>% 
  gghistogram(x = "rol")


h9 = adm.imp %>% 
  gghistogram(x = "rq")


h10 = adm.imp %>% 
  gghistogram(x = "polstab")

h11 = adm.imp %>% 
  gghistogram(x = "voice")

## Visualize 
ggarrange(h6, h7, h8, h9, h10, h11)

```


## **Random Forest**

```{r Adm1 RF iml}
# load libraries
library(randomForest) #rf package
library(iml) #interpretable machine learning package

```

```{r Adm1 RF Data Frame}
# subset the variables we want to use for this model
# remove zscore and adm1 (adm1 has too many factors)
adm.imp.rf.df = adm.imp %>% 
  select(-zscore, -adm1)

str(adm.imp.rf.df)

```

```{r Adm 1 RF Train}
st = Sys.time()

# Train the random forest 
set.seed(123)
rf = randomForest(con ~ .,
                  data = adm.imp.rf.df)

et = Sys.time()
elapsed = et-st
print(elapsed)
```

```{r Adm1 RF Errors}
##pseudo r2 value
mean(rf$rsq)

```

### *iml Objects and Interpretations**

```{r Adm1 Predictor Object}
## Create a data frame with the features minus the target
x = adm.imp.rf.df %>% 
  select(-con)

# Store the data and con in the predictor container along with the randomForest parameters
adm.imp.predictor = Predictor$new(model = rf,
                                  data = x,
                                  y = adm.imp.rf.df$con)

```


```{r Adm1 Feature Importance, message=FALSE}
## Store the features in a FeatureImp object
## loss argument specifies the performance measure for error
importance = FeatureImp$new(predictor = adm.imp.predictor,
                     loss = "rmse")

```


```{r Adm1 Importance Visualization}
## Visualize the importance using ggplot2
library(ggplot2)

##Plot
plot(importance)

```

**Feature Importance**

```{r Adm1 Importance Results}
## View the the feature importance percentiles
importance$results

```


**Feature Effects**

```{r Adm1 ALE Plot}
## Reminder this allows us to interpret areas of the curve that may be more important than others
##grid.size argument is the number of quantiles specified

ale = FeatureEffect$new(predictor = adm.imp.predictor,
                        feature = "pop",
                        grid.size = 10)

ale$plot()
```

```{r Adm1 ALE Plot Loop}
## Define a helper function to plot

ale.plot = function(x){
  ale$set.feature(x)
  ale$plot()
}

```

```{r Adm1 ALE Plots Complete}
# plot through the variables
ale.plot("country")
ale.plot("temp")
ale.plot("polstab")
ale.plot("precip")
ale.plot("voice")
ale.plot("ge")
ale.plot("cpt")
ale.plot("nlights")
ale.plot("rol")
ale.plot("rq")
ale.plot("year")
ale.plot("pop")

```

```{r Adm 1 Plot Feature Effects - All}
##Plot all of the feature effects at once

effs = FeatureEffects$new(predictor = adm.imp.predictor,
                          grid.size = 10)

# Visualize
plot(effs)

```

**Partial Dependecy Plots**

**Note:** "Individual Conditional Expectation (ICE) plots display one line per instance that shows how the instance’s prediction changes when a feature changes.

The partial dependence plot for the average effect of a feature is a global method because it does not focus on specific instances, but on an overall average. The equivalent to a PDP for individual data instances is called individual conditional expectation (ICE) plot (Goldstein et al. 201748)."

The yellow line represents the partial dependency of the feature, which is the average of the ICE plots.

```{r Adm1 PDPs, message=FALSE, warning=FALSE}
#population
#store it in an object
rf.pop = Partial$new(predictor = adm.imp.predictor,
                        feature = "pop",
                        aggregation = "pdp",
                        ice = TRUE)

rf.country = Partial$new(predictor = adm.imp.predictor,
                        feature = "country",
                        aggregation = "pdp",
                        ice = TRUE)

rf.temp = Partial$new(predictor = adm.imp.predictor,
                        feature = "temp",
                        aggregation = "pdp",
                        ice = TRUE)

rf.precip = Partial$new(predictor = adm.imp.predictor,
                        feature = "precip",
                        aggregation = "pdp",
                        ice = TRUE)

rf.nlights = Partial$new(predictor = adm.imp.predictor,
                        feature = "nlights",
                        aggregation = "pdp",
                        ice = TRUE)

rf.ge = Partial$new(predictor = adm.imp.predictor,
                        feature = "ge",
                        aggregation = "pdp",
                        ice = TRUE)

rf.cpt = Partial$new(predictor = adm.imp.predictor,
                        feature = "cpt",
                        aggregation = "pdp",
                        ice = TRUE)

rf.rol = Partial$new(predictor = adm.imp.predictor,
                        feature = "rol",
                        aggregation = "pdp",
                        ice = TRUE)

rf.rq = Partial$new(predictor = adm.imp.predictor,
                        feature = "rq",
                        aggregation = "pdp",
                        ice = TRUE)

rf.polstab = Partial$new(predictor = adm.imp.predictor,
                        feature = "polstab",
                        aggregation = "pdp",
                        ice = TRUE)

rf.voice = Partial$new(predictor = adm.imp.predictor,
                        feature = "voice",
                        aggregation = "pdp",
                        ice = TRUE)

# center (this centers the impact of y hat on a starting value)
rf.pop$center(min(adm.imp.rf.df$pop))
rf.temp$center(min(adm.imp.rf.df$temp))
rf.precip$center(min(adm.imp.rf.df$precip))
rf.nlights$center(min(adm.imp.rf.df$nlights))
rf.ge$center(min(adm.imp.rf.df$ge))
rf.cpt$center(min(adm.imp.rf.df$cpt))
rf.rol$center(min(adm.imp.rf.df$rol))
rf.rq$center(min(adm.imp.rf.df$rq))
rf.polstab$center(min(adm.imp.rf.df$polstab))
rf.voice$center(min(adm.imp.rf.df$voice))


```


```{r Adm1 PDPs 2}
# plot
p1 = plot(rf.pop) + ggtitle("Population PDP/ICE Plot")
p2 = plot(rf.country) + ggtitle("Country PDP/ICE Plot")
p3 = plot(rf.temp) + ggtitle("Temperature PDP/ICE Plot")
p4 = plot(rf.precip) + ggtitle("Precipitation PDP/ICE Plot")

gridExtra::grid.arrange(p1, p2, p3, p4, nrow = 2)

```


```{r Adm1 PDPs 3}
# plot
p5 = plot(rf.nlights) + ggtitle("Nighttime Lights PDP/ICE Plot")
p6 = plot(rf.ge) + ggtitle("Governance Effectiveness PDP/ICE Plot")
p7 = plot(rf.cpt) + ggtitle("Corruption PDP/ICE Plot")
p8 = plot(rf.rol) + ggtitle("Rule of Law PDP/ICE Plot")

gridExtra::grid.arrange(p5, p6, p7, p8, nrow = 2)

```

```{r Adm1 PDPs 4}
# plot
p9 = plot(rf.rq) + ggtitle("Regulatory Quality PDP/ICE Plot")
p10 = plot(rf.polstab) + ggtitle("Political Stability PDP/ICE Plot")
p11 = plot(rf.voice) + ggtitle("Voice and Accountability PDP/ICE Plot")


gridExtra::grid.arrange(p9, p10, p11, nrow = 2)

```


**Measure Interactions**

```{r Adm 1 Interactions}
st = Sys.time()
##Set up the interactions wrapper
##Play around with grid.size

adm.interact = Interaction$new(predictor = adm.imp.predictor,
                           grid.size = 15)

et = Sys.time()
print(et-st)
```

```{r Adm 1 Plot Interactions}
##Plot the features to see how the interact with any other feature in the data
# scale 0 - 1 with 1 meaning that 100% of the variance is explained with interactions with the other features
plot(adm.interact)

```


## **Surrogate Model**

```{r Adm 1 Surrogate Model}

tree = TreeSurrogate$new(predictor = adm.imp.predictor,
                         maxdepth = 2)


# Visualize
plot(tree)

```


```{r Adm 1 Plot The RF Tree}

# Use this code to plot the random forest tree itself
plot(tree$tree)

```


```{r Adm 1 Surrogate Model Evaluation} 
## Check the model performance 

print(tree$r.squared)

# 0.47 - doesn't perform that well

```





# **Administrative Level 1 Analysis (Non-Imputed Data)**

This analysis re-runs the model code on the non-imputed data frame to see if there is any noticeable differences in performance or outcomes.


```{r Non-imp data read}
## read in the non-imputed data
adm.nonimp = read.csv("../cleaned_data/hma_adm1_non_imp_df_clean_20240418.csv")

# remove the "X" column
adm.nonimp = adm.nonimp %>% 
  select(-X)

# factor country and adm 1
adm.nonimp$country = as.factor(adm.nonimp$country)
adm.nonimp$adm1 = as.factor(adm.nonimp$adm1)

# check 
str(adm.nonimp)

```

## **Random Forest (Non-Imputed Data)**


```{r Non-Imp RF Data Frame}
# subset the variables we want to use for this model
# remove adm1 (adm1 has too many factors)
adm.nonimp.rf.df = adm.nonimp %>% 
  select(-adm1)

str(adm.nonimp.rf.df)

```

```{r Non-Imp RF Train}
st = Sys.time()

# Train the random forest 
set.seed(456)
nonimp.rf = randomForest(con ~ .,
                  data = adm.nonimp.rf.df)

et = Sys.time()
elapsed = et-st
print(elapsed)
```

```{r Non-Imp RF Errors}
##pseudo r2 value
mean(nonimp.rf$rsq)

```

```{r Performance Comparison}
# Imputed model performance
mod1.rsq = mean(rf$rsq)

# Non-imputed model performance
mod2.rsq = mean(nonimp.rf$rsq)

print(c(mod1.rsq, mod2.rsq))

```

### **Non-Imp iml Objects and Interpretations**

```{r Non Imp Predictor Object}
## Create a data frame with the features minus the target
y = adm.nonimp.rf.df %>% 
  select(-con)

# Store the data and con in the predictor container along with the randomForest parameters
adm.nonimp.predictor = Predictor$new(model = nonimp.rf,
                                  data = y,
                                  y = adm.nonimp.rf.df$con)

```

**Feature Importance**

```{r Non Imp Feature Importance, message=FALSE}
## Store the features in a FeatureImp object
## loss argument specifies the performance measure for error
nonimp.importance = FeatureImp$new(predictor = adm.nonimp.predictor,
                     loss = "rmse")

```


```{r Non Imp Importance Visualization}

##Plot
plot(nonimp.importance)

```

**ALE Plots**

```{r Non Imp Plot Feature Effects - All}
##Plot all of the feature effects at once

nonimp.effs = FeatureEffects$new(predictor = adm.nonimp.predictor,
                          grid.size = 10)

# Visualize
plot(nonimp.effs)

```


**Measure Interactions**

```{r Non Imp Interactions}
st = Sys.time()
##Set up the interactions wrapper
##Play around with grid.size

nonimp.interact = Interaction$new(predictor = adm.nonimp.predictor,
                           grid.size = 15)

et = Sys.time()
print(et-st)
```

```{r Non Imp Plot Interactions}
##Plot the features to see how the interact with any other feature in the data
# scale 0 - 1 with 1 meaning that 100% of the variance is explained with interactions with the other features
plot(nonimp.interact)

```


### **Surrogate Model**

```{r Non Imp Surrogate Model}

tree.nonimp = TreeSurrogate$new(predictor = adm.nonimp.predictor,
                         maxdepth = 2)


# Visualize
plot(tree.nonimp)

```

```{r Non Imp Surrogate Tree Viz}
# visualize the simplified tree

plot(tree.nonimp$tree)

```

```{r Non Imp Surrogate R Squared}
# How well does the surrogate model do at predicting conflict
print(tree.nonimp$r.squared)

```



# **Country Level Analysis (Imputed Data)**

This analysis was performed on a dataframe that was imputed using a mean-by-country approach.

```{r Country Imp Data}
# read in the data
country.imp = read.csv("../cleaned_data/hma_country_df_clean_20240418.csv")

str(country.imp)

country.imp = country.imp |>
  select(-X) |>
  mutate(country = as.factor(country))

# check
str(country.imp)

```

## **Random Forest**

```{r Country Imp RF}

st = Sys.time()

# Train the random forest 
set.seed(923)
country.imp.rf = randomForest(con ~ ., data = country.imp)

et = Sys.time()
elapsed = et-st
print(elapsed)

```

```{r Country Imp R Squared}
## pseudo r2 value
mean(country.imp.rf$rsq)

## RMSE 
c.imp.rmse = sqrt(mean(country.imp.rf$mse))

## percentage error

c.imp.rmse / diff(range(country.imp$con))

##*NOTE* Followup with Simon/Andrew on this -
##*does using the difference instead of mean conflict count mean we will likely over predict low conflict areas???

# R2 is great..mean calculation of error is not so great...calculating it on the range is good.
```

## **iml Predictor**

```{r Country Imp Predictor DF}
## Create a data frame with the features minus the target
country.imp.pred.df = country.imp %>% 
  select(-con)

```


```{r Country Imp Predictor}
# Store the data and con in the predictor container along with the randomForest parameters
country.imp.predictor = Predictor$new(model = country.imp.rf,
                                      data = country.imp.pred.df,
                                      y = country.imp$con)
```


```{r Country Imp Importance Object}
## Store the features in a FeatureImp object
## loss argument specifies the performance measure for error
country.imp.importance = FeatureImp$new(predictor = country.imp.predictor,
                                        loss = "rmse")

```


```{r Country Imp Importance Plot}
##Plot
plot(country.imp.importance)

```


```{r Country Imp ALE Plots}
country.imp.ale = FeatureEffect$new(predictor = country.imp.predictor,
                                    feature = "rol",
                                    grid.size = 10)

country.imp.ale$plot()

```

```{r Country Imp ALE Effects Plot ALL}
##Plot all of the feature effects at once

country.imp.effs = FeatureEffects$new(predictor = country.imp.predictor,
                                      grid.size = 10)

# Visualize
plot(country.imp.effs)
```


```{r Country Imp PDPs, warning=FALSE}

#PDPs
country.imp.rf.rol = Partial$new(predictor = country.imp.predictor,
                     feature = "rol",
                     aggregation = "pdp",
                     ice = TRUE)

country.imp.rf.country = Partial$new(predictor = country.imp.predictor,
                     feature = "country",
                     aggregation = "pdp",
                     ice = TRUE)

country.imp.rf.polstab = Partial$new(predictor = country.imp.predictor,
                     feature = "polstab",
                     aggregation = "pdp",
                     ice = TRUE)

country.imp.rf.pop = Partial$new(predictor = country.imp.predictor,
                     feature = "pop",
                     aggregation = "pdp",
                     ice = TRUE)

country.imp.rf.cd = Partial$new(predictor = country.imp.predictor,
                     feature = "cd",
                     aggregation = "pdp",
                     ice = TRUE)

country.imp.rf.voice = Partial$new(predictor = country.imp.predictor,
                     feature = "voice",
                     aggregation = "pdp",
                     ice = TRUE)


# center (this centers the impact of y hat on a starting value)
country.imp.rf.rol$center(min(country.imp$rol))
country.imp.rf.polstab$center(min(country.imp$polstab))
country.imp.rf.pop$center(min(country.imp$pop))
country.imp.rf.cd$center(min(country.imp$cd))
country.imp.rf.voice$center(min(country.imp$voice))



```

```{r Country Imp PDP Plots All}
library(gridExtra)

p1 = plot(country.imp.rf.rol) + ggtitle("Rule of Law PDP/ICE Plot")
p2 = plot(country.imp.rf.country) + ggtitle("Country PDP/ICE Plot")
p3 = plot(country.imp.rf.polstab) + ggtitle("Political Stability PDP/ICE Plot")
p4 = plot(country.imp.rf.pop) + ggtitle("Population PDP/ICE Plot")
p5 = plot(country.imp.rf.cd) + ggtitle("Climate Disasters PDP/ICE Plot")
p6 = plot(country.imp.rf.voice) + ggtitle("Voice PDP/ICE Plot")

grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 2)
```



```{r Country Imp Interaction}
country.imp.interact = Interaction$new(predictor = country.imp.predictor,
                               grid.size = 15)

et = Sys.time()
print(et-st)

##Plot the features to see how the interact with any other feature in the data
# scale 0 - 1 with 1 meaning that 100% of the variance is explained with interactions with the other features
plot(country.imp.interact)

```
## **Surrogate Model**

```{r Country Imp Surrogate}
# Surrogate Model

country.imp.tree = TreeSurrogate$new(predictor = country.imp.predictor,
                                      maxdepth = 2)


# Visualize
plot(country.imp.tree)

```


```{r Country Imp Surrogate Visualization}

# Visualize Simplified version
plot(country.imp.tree$tree)

```


```{r Country Imp Surrogate Performance}

## Check the model performance 

print(country.imp.tree$r.squared)

```





# **Country Level Analysis (Imputed with missForest)**

This analysis was performed using a data frame imputed using missForest.


```{r mf Country Imp Data}
# read in the data
mf.country.imp = read.csv("../cleaned_data/hma_country_MF_impute_df_clean_20240424.csv")

str(mf.country.imp)

mf.country.imp = mf.country.imp |>
  select(-X) |>
  mutate(country = as.factor(country))

# check
str(mf.country.imp)

```

## **Random Forest**

```{r mf Country Imp RF}

st = Sys.time()

# Train the random forest 
set.seed(764)
mf.country.imp.rf = randomForest(con ~ ., data = mf.country.imp)

et = Sys.time()
elapsed = et-st
print(elapsed)

```

```{r mf Country Imp R Squared}
## pseudo r2 value
mean(mf.country.imp.rf$rsq)

## RMSE 
mf.c.imp.rmse = sqrt(mean(mf.country.imp.rf$mse))

## percentage error

mf.c.imp.rmse / diff(range(mf.country.imp$con))

##*NOTE* Followup with Simon/Andrew on this -
##*does using the difference instead of mean conflict count mean we will likely over predict low conflict areas???

# R2 is great..mean calculation of error is not so great...calculating it on the range is good.
```

## **iml Predictor**

```{r mf Country Imp Predictor DF}
## Create a data frame with the features minus the target
mf.country.imp.pred.df = mf.country.imp %>% 
  select(-con)

```


```{r mf Country Imp Predictor}
# Store the data and con in the predictor container along with the randomForest parameters
mf.country.imp.predictor = Predictor$new(model = mf.country.imp.rf,
                                      data = mf.country.imp.pred.df,
                                      y = mf.country.imp$con)
```


```{r mf Country Imp Importance Object}
## Store the features in a FeatureImp object
## loss argument specifies the performance measure for error
mf.country.imp.importance = FeatureImp$new(predictor = mf.country.imp.predictor,
                                        loss = "rmse")

```


```{r mf Country Imp Importance Plot}
##Plot
plot(mf.country.imp.importance)

```


```{r mf Country Imp ALE Plots}
mf.country.imp.ale = FeatureEffect$new(predictor = mf.country.imp.predictor,
                                    feature = "rol",
                                    grid.size = 10)

mf.country.imp.ale$plot()

```

```{r mf Country Imp ALE Effects Plot ALL}
##Plot all of the feature effects at once

mf.country.imp.effs = FeatureEffects$new(predictor = mf.country.imp.predictor,
                                      grid.size = 10)

# Visualize
plot(mf.country.imp.effs)
```


```{r mf Country Imp PDPs, warning=FALSE}

#PDPs
mf.country.imp.rf.rol = Partial$new(predictor = mf.country.imp.predictor,
                     feature = "rol",
                     aggregation = "pdp",
                     ice = TRUE)

mf.country.imp.rf.country = Partial$new(predictor = mf.country.imp.predictor,
                     feature = "country",
                     aggregation = "pdp",
                     ice = TRUE)

mf.country.imp.rf.polstab = Partial$new(predictor = mf.country.imp.predictor,
                     feature = "polstab",
                     aggregation = "pdp",
                     ice = TRUE)

mf.country.imp.rf.pop = Partial$new(predictor = mf.country.imp.predictor,
                     feature = "pop",
                     aggregation = "pdp",
                     ice = TRUE)

mf.country.imp.rf.cd = Partial$new(predictor = mf.country.imp.predictor,
                     feature = "cd",
                     aggregation = "pdp",
                     ice = TRUE)

mf.country.imp.rf.voice = Partial$new(predictor = mf.country.imp.predictor,
                     feature = "voice",
                     aggregation = "pdp",
                     ice = TRUE)


# center (this centers the impact of y hat on a starting value)
mf.country.imp.rf.rol$center(min(mf.country.imp$rol))
mf.country.imp.rf.polstab$center(min(mf.country.imp$polstab))
mf.country.imp.rf.pop$center(min(mf.country.imp$pop))
mf.country.imp.rf.cd$center(min(mf.country.imp$cd))
mf.country.imp.rf.voice$center(min(mf.country.imp$voice))



```

```{r mf Country Imp PDP Plots All}
library(gridExtra)

mf.p1 = plot(mf.country.imp.rf.rol) + ggtitle("Rule of Law PDP/ICE Plot")
mf.p2 = plot(mf.country.imp.rf.country) + ggtitle("Country PDP/ICE Plot")
mf.p3 = plot(mf.country.imp.rf.polstab) + ggtitle("Political Stability PDP/ICE Plot")
mf.p4 = plot(mf.country.imp.rf.pop) + ggtitle("Population PDP/ICE Plot")
mf.p5 = plot(mf.country.imp.rf.cd) + ggtitle("Climate Disasters PDP/ICE Plot")
mf.p6 = plot(mf.country.imp.rf.voice) + ggtitle("Voice PDP/ICE Plot")

grid.arrange(mf.p1, mf.p2, mf.p3, mf.p4, mf.p5, mf.p6, nrow = 2)
```



```{r mf Country Imp Interaction}
mf.country.imp.interact = Interaction$new(predictor = mf.country.imp.predictor,
                               grid.size = 15)

et = Sys.time()
print(et-st)

##Plot the features to see how the interact with any other feature in the data
# scale 0 - 1 with 1 meaning that 100% of the variance is explained with interactions with the other features
plot(mf.country.imp.interact)

```
## **Surrogate Model**

```{r mf Country Imp Surrogate}
# Surrogate Model

mf.country.imp.tree = TreeSurrogate$new(predictor = mf.country.imp.predictor,
                                      maxdepth = 2)


# Visualize
plot(mf.country.imp.tree)

```


```{r mf Country Imp Surrogate Visualization}

# Visualize Simplified version
plot(mf.country.imp.tree$tree)

```


```{r mf Country Imp Surrogate Performance}

## Check the model performance 

print(mf.country.imp.tree$r.squared)

```



# **Country Level Analysis (Non-Imputed Data)**

```{r Country Non Imp}
# Read in the clean data frame
country.nonimp = read.csv("../cleaned_data/hma_country_non_imp_df_clean_20240418.csv")

# Remove X and change country to a factor
country.nonimp = country.nonimp %>% 
  select(-X) %>% 
  mutate(country = as.factor(country))

# check
summary(country.nonimp)

```

Note: We need to remove the NAs from the dataframe before running the random forest. I fixed the orignal csv file that accidentally introduced NAs into the dataframe. 

```{r Country Non Imp NA Drop}
#drop_na from tidyr to filter out columns with nas 

#country.nonimp.cleaned = country.nonimp |>
  #drop_na()


# check
#summary(country.nonimp.cleaned)

```



## **Random Forest**

```{r Country Non Imp RF}

st = Sys.time()

# Train the random forest 
set.seed(429)
country.nonimp.rf = randomForest(con ~ ., data = country.nonimp)

et = Sys.time()
elapsed = et-st
print(elapsed)

```

```{r Country Non Imp R Squared}
## pseudo r2 value
mean(country.nonimp.rf$rsq)

## RMSE 
country.nonimp.rmse = sqrt(mean(country.nonimp.rf$mse))

## percentage error

country.nonimp.rmse / diff(range(country.nonimp$con))

##*NOTE* Followup with Simon/Andrew on this -
##*does using the difference instead of mean conflict count mean we will likely over predict low conflict areas???

# R2 is great..mean calculation of error is not so great...calculating it on the range is good.
```

## **iml Predictor**

```{r Country Non Imp Predictor DF}
## Create a data frame with the features minus the target
country.nonimp.pred.df = country.nonimp %>% 
  select(-con)

```


```{r Country NonImp Predictor}
# Store the data and con in the predictor container along with the randomForest parameters
country.nonimp.predictor = Predictor$new(model = country.nonimp.rf,
                                      data = country.nonimp.pred.df,
                                      y = country.nonimp$con)
```


```{r Country Non Imp Importance Object}
## Store the features in a FeatureImp object
## loss argument specifies the performance measure for error
country.nonimp.importance = FeatureImp$new(predictor = country.nonimp.predictor,
                                        loss = "rmse")

```


```{r Country Non Imp Importance Plot}
##Plot
plot(country.nonimp.importance)

```


## **Surrogate Model**

```{r Country NonImp Surrogate}
# Surrogate Model

country.nonimp.tree = TreeSurrogate$new(predictor = country.nonimp.predictor,
                                      maxdepth = 2)


# Visualize
plot(country.nonimp.tree)

```


```{r Country NonImp Surrogate Visualization}

# Visualize Simplified version
plot(country.nonimp.tree$tree)

```


```{r Country NonImp Surrogate Performance}

## Check the model performance 

print(country.nonimp.tree$r.squared)

```



# **Figures and General Summary Statistics**

```{r Plot by Country}
## Visualize using ggplot
## Reminder - i used the average conflict to negate 0 values for years with no data

con.plot = ggplot(data = country.nonimp, aes(x = year, y = con, color = country)) +
  geom_point() +
  geom_line(lwd = 1.2) +
  geom_smooth(method = "lm",
              se = FALSE) +
  facet_wrap(~country, scales = "free") +
  scale_x_continuous(breaks = c(seq(from = 2010, to = 2022, by = 1))) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5)) +
  labs(title = "Total Conflict Count (2010-2022)",
       color = "Country") +
  xlab("Year") +
  ylab("Conflict Count")
  

## View
con.plot
```







## **Summary Statistics**

```{r Summary Statistics - Country}

country.summary = country.imp %>% 
  summarize(sd.con = sd(con),
            sd.gdp = sd(gdp),
            sd.pop = sd(pop),
            sd.temp = sd(temp),
            sd.cd = sd(cd),
            sd.ge = sd(ge),
            sd.cpt = sd(cpt),
            sd.rol = sd(rol),
            sd.rq = sd(rq),
            sd.pol = sd(polstab),
            sd.voice = sd(voice),
            sd.urban = sd(urbanpop),
            sd.water = sd(waterwith))


#print
country.summary

```


```{r Summary Statistics - ADM1}

adm1.summary = adm.imp %>% 
  summarize(sd.nlights = sd(nlights),
            sd.pop = sd(pop),
            sd.temp = sd(temp),
            sd.precip = sd(precip),
            sd.con = sd(con))

#print
print(adm1.summary)

```














