---
title: "Pakistan Water Conflict Data Cleaning Pt 2"
author: "David Leydet"
date: "2024-10-09"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: 
      collapsed: false
      smooth_scroll: false
    theme: bootstrap
    df_print: paged
   
---


This step takes our wide data frame (pak_water_complete_df_20241007_v2), converts it to long form, and imputes missing variables using missForest.


```{r}
# setwd

setwd("../r_code/")

```


# **Initial Data Read and Pivot**


```{r}

wide = read.csv("../data/pak_water_2024_run/pak_water_complete_df_20241007_v2.csv")

str(wide)

```


```{r}
library(tidyverse)

# convert each column (except the first one - admin2) to numeric

wide[ ,2:310] <- sapply(wide[,2:310], as.numeric)

#check
str(wide)



```




```{r}
# pivot to long data frame

long = wide %>% 
  pivot_longer(cols = -c(admin2),
    names_sep = "\\.",
    names_to = c(".value", "year")
  )

#check
str(long)

#summary
summary(long)

#visual check
#write.csv(long, file = "~/Desktop/pak_water_long_check.csv")

```


Think about removing the Indian Occupied Kashmir observations. There are no conflict observations in the ACLED data for Pakistan in this area. It is likely coded for India. 

Run two versions of the model?


# **missForest Imputation**


This section of code imputes variables with NAs using missForest.

The following imputation uses the missForest R package. Details on the imputation procedures can be found [here](https://cran.r-project.org/web/packages/missForest/missForest.pdf).

**Note:** missForest requires the use of a data frame object (not a tibble), factors (not characters) for non-numeric data, and cannot impute with a large number of classes. In this case we change country, adm1, and year in to factors or numerics. We subsequently removed the adm1 from the data frame for imputation, then return it afterwards.

Additionally, the missForest object stores two things:
1. ximp - this contains the imputed data frame.
2. OOB Error - this is the Normalized RMSE error associated with the imputation.


```{r}

# create a new dataframe
long.mf = long %>% 
  mutate(admin2 = as.factor(admin2)) %>% # missForest requires admin2 to be a factor
  mutate(year = as.numeric(year)) # needs to be numeric not character

# convert from tibble to dataframe
long.mf = as.data.frame(long.mf)

# write this data frame (to use for modeling)
#write.csv(long.mf, file = "../data/pak_water_2024_run/pak_water_long_nonimp_20241010.csv")

# Note - every row (observation) has some NAs.

# remove the admin2 for missForest to run correctly. We will add this back after the imputation
long.mf.no.admin2 = long.mf %>% 
  select(-admin2)

#check
str(long.mf.no.admin2)

```



```{r}
# run missForest
library(missForest)

long.impute = missForest(long.mf.no.admin2)

# save the dataframe
long.impute.df = long.impute$ximp



```

## **Check Imputation**

```{r}
# OOB error
long.impute$OOBerror


```



```{r}
##Check the imputation to see how well it did
##nlights
plot(density(long$landscan_global_population, na.rm=TRUE))
lines(density(long.impute$ximp$landscan_global_population), col = "red", lty = 2)

##conflict
plot(density(long$total, na.rm=TRUE))
lines(density(long.impute$ximp$total), col = "red", lty = 2)

```

```{r}

density.plot = function(x){
  plot(density(long$x, na.rm=TRUE))
lines(density(long.impute$ximp$x), col = "red", lty = 2)
}

lapply(names(long.mf.no.admin2), density.plot)

for (i in colnames(long.mf.no.admin2)){
  plot(density(long$[[i]], na.rm=TRUE))
  lines(density(long.impute$ximp$[[i]]), col = "red", lty = 2)
}

```


```{r}
library(ggplot2)

long_imp = long.impute$ximp

# create a long df for the imputed data
long_imp_long <- long_imp |>
  pivot_longer(cols = names(long_imp), names_to = "var") |>
  mutate(imputed = "1")

# create a long df for the raw data
long_raw_long <- long.mf |>
  select(!admin2) |>
  pivot_longer(cols = names(long.mf[, -1]), names_to = "var") |> # long.mf[,-1] removes the admin2 row
  mutate(imputed = "0")

long_combined <- rbind(long_imp_long, long_raw_long)

ggplot(long_combined, aes(value, colour = imputed)) +
  geom_density() +
  facet_wrap(~var, scales = "free")


```


## **New Complete Dataframe**

```{r}
## add back in admin2

long.impute.df$admin2 = long$admin2

# relocate admin2 before year
long.impute.df = long.impute.df %>% 
  relocate(admin2, .before = year)

# check
str(long.impute.df)

# write the df
#write.csv(long.impute.df, file = "../data/pak_water_2024_run/pak_water_imputed_df_full_20241010.csv")

```




















