---
title: "Pakistan Water Conflict"
author: "David Leydet"
date: "2024-06-26"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: 
      collapsed: false
      smooth_scroll: false
    theme: bootstrap
    df_print: paged
   
---


# **Set up**


```{r Working Dir, message=FALSE, warning=FALSE}
# set working directory

setwd("~/Desktop/University of Utah PhD /Research/r_code/")

# load the usual libraries
library(tidyverse) #core packages of tidyverse
library(janitor) # count/summary package

```



```{r ACLED Data Load}
# ACLED Data Load
# This is the complete Pakistan ACLED data from January 1, 2010 - December 31, 2023
pak = read.csv("../data/pak_water_2024_run/pak_con_overall_2010_2023.csv")

# check 
str(pak)

```


```{r}
## Total Conflict Counts by district
library(janitor)

pak.counts = janitor::tabyl(dat = pak, admin2, year)

pak.counts

```


```{r}
# rename column names
col.name.change = c(admin2 = "admin2", total.2010 = "2010", total.2011 = "2011", total.2012 = "2012", total.2013 = "2013", total.2014 = "2014", total.2015 = "2015", total.2016 = "2016", total.2017 = "2017", total.2018 = "2018", total.2019 = "2019", total.2020 = "2020", total.2021 = "2021", total.2022 = "2022", total.2023 = "2023")

# change the column names and remove the admin2 row that is blank (these are sea-based conflicts)
pak.counts.2 = pak.counts %>% 
  rename(all_of(col.name.change)) %>% 
  dplyr::filter(admin2 !="")

str(pak.counts.2)

```



```{r Extract Water-related conflicts}

## Extract Water Events (Wolf et al., 2003 methodology for key terms)

water = dplyr::filter(pak, grepl("water|flood|^rain|^river|snow|lake|^dam|stream|tributary|irrigation|diversion|drought|channel|canal|fish|resorvoir|hydroelect", pak$notes, ignore.case = TRUE))

# visual check (complete)
#view(water)

# NA check
#na_counts = water |>
  #summarise_all( ~ sum(is.na(.)))

#print(na_counts)

```

```{r}
# Identify the NAs

blank = water %>% 
  dplyr::filter(admin2 == "")

print(blank)

```

```{r write water csv}
## remove the NAs (see notes about the 6 observations removed - all were sea-based disputes with India)
# start with 4763 obs

water = water %>% 
  dplyr::filter(admin2 !="")

str(water)

# end with 4757 obs

# write csv

#write.csv(water, file = "../data/pak_water_2024_run/pak_water_con_total_2010_2023_20240911.csv")

```



```{r Count by district}
# count by district

#water %>% 
  #group_by(admin2) %>% 
  #count(year) %>% 
  #ungroup() %>% 
  

```


```{r Counts by district data frame}
# create a new data frame with the counts by district

#dis.counts = water %>% 
  #group_by(admin2) %>% 
  #count(year) %>% 
  #ungroup() %>% 
  #dplyr::filter(admin2 != "") # remove the NAs (see notes about the 6 observations removed - all were sea-based disputes with India)

#print(dis.counts)

```

```{r write district counts}
# write the district counts to a csv file

#write.csv(dis.counts, file = "../data/pak_water_2024_run/pak_water_con_district_counts_20240911.csv")


```


```{r District Names}
# test vector of the district names to ensure they match with the GeoQuery admin 2 spatial units

#adm2 = unique(water$admin2)

```



# **Pak Water Excel Data Cleaning**

The following code cleans up the Pakistan Water Conflict Total file. We need to count the conflicts by district from 2010-2023. Then, we need to merge this file with our GeoQuery covariates to complete the data frame for analysis. 


```{r}
# Ensure you read in the pak_water_con_total_2010_2023_2024_0911.csv file
# This is the subset of the overall ACLED data 


df = read.csv("../data/pak_water_2024_run/pak_water_con_total_2010_2023_20240911.csv")


```


```{r}
# counts using janitor
# district data frame counts by year

library(janitor)

dis.df = janitor::tabyl(dat = df, admin2, year) 


# check
str(dis.df)


```


```{r}

# write the dis.df to csv
# combine this with the pak_geoquery_20241006_preprocess.csv file

#write.csv(dis.df, file = "../data/pak_water_2024_run/district_counts_20241007.csv")

```


## **Merge dataframes**

Merge the pak_geoquery_20241006_preprocess.csv file and the dis.df dataframe. This combines the variables from geoquery with the ACLED water conflict counts. 

```{r}

# rename column names
water.col.name.change = c(admin2 = "admin2", water.2010 = "2010", water.2011 = "2011", water.2012 = "2012", water.2013 = "2013", water.2014 = "2014", water.2015 = "2015", water.2016 = "2016", water.2017 = "2017", water.2018 = "2018", water.2019 = "2019", water.2020 = "2020", water.2021 = "2021", water.2022 = "2022", water.2023 = "2023")

# change the column names and remove the admin2 row that is blank (these are sea-based conflicts)
dis.df.2 = dis.df %>% 
  rename(all_of(water.col.name.change)) %>% 
  dplyr::filter(admin2 !="")

str(dis.df.2)

```


```{r}
# merge the pak.counts.2 (total conflict counts) along with the water conflict dataframe

con.counts.full.df = full_join(x = dis.df.2, y = pak.counts.2, by = "admin2")

view(con.counts.full.df)

```


```{r}
# add percentage column to 

con.counts.full.df = con.counts.full.df %>% 
  mutate(perc.2010 = (water.2010/total.2010)*100) %>% 
  mutate(perc.2011 = (water.2011/total.2011)*100) %>% 
  mutate(perc.2012 = (water.2012/total.2012)*100) %>% 
  mutate(perc.2013 = (water.2013/total.2013)*100) %>% 
  mutate(perc.2014 = (water.2014/total.2014)*100) %>% 
  mutate(perc.2015 = (water.2015/total.2015)*100) %>% 
  mutate(perc.2016 = (water.2016/total.2016)*100) %>% 
  mutate(perc.2017 = (water.2017/total.2017)*100) %>% 
  mutate(perc.2018 = (water.2018/total.2018)*100) %>% 
  mutate(perc.2019 = (water.2019/total.2019)*100) %>% 
  mutate(perc.2020 = (water.2020/total.2020)*100) %>% 
  mutate(perc.2021 = (water.2021/total.2021)*100) %>% 
  mutate(perc.2022 = (water.2022/total.2022)*100) %>% 
  mutate(perc.2023 = (water.2023/total.2023)*100) 
  

# replace NaNs with 0
# NaNs were generated by a divisor - in this case the total number of conflicts - of zero

con.counts.full.df[is.na(con.counts.full.df)] = 0

summary(con.counts.full.df)

```

```{r}
#write the combined file

#write.csv(con.counts.full.df, file = "../data/pak_water_2024_run/conflict_full_df_20231007.csv" )

```






