---
title: "Pakistan Conflict Modeling with iml Package"
author: "David Leydet"
date: "2025-06-05"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: 
      collapsed: false
      smooth_scroll: false
    theme: bootstrap
    df_print: paged
   
---


This file runs the Pakistan conflict data through a random forest using the iml package. This was completed to address reviewer comments from the Weather, Climate, and Society journal.  



# **Data**

```{r}

#set working directory

setwd("../r_code/")


```


```{r}

library(dplyr)

# read in data

data = read.csv("../cleaned_data/pakistan_conflict_data_20250604.csv")

# remove the X column

data = data |>
  dplyr::select(-c(X, log.total.con.percapita)) %>% 
  mutate(log.con.percapita = log(((data$total.con+1) / data$pop)*10000)) %>% #the model target is this variable!
  mutate(total.con.pertenthousand = data$total.con.percapita*10000) #per 10000 variable created for future work

# check
str(data)

```

# **Data Viz**

```{r}
# percentage of protests to overall conflict

(sum(data$protests)/sum(data$total.con))*100

```



# **Scatterplot**

```{r}
library(ggplot2)

ggplot(data, aes(x=year, y = log.con.percapita)) +
  geom_point() +
  geom_smooth()

```


# **Model**

```{r, messages=FALSE}
# load iml and random forest packages

library(iml)
library(randomForest)

```



```{r}

# subset the data to include the variables that you want to examine 

variables = c("districts", "year", "spei", "pop", "dep.index", "built", "temp.jan", "temp.apr", "temp.jul", "temp.oct", "precip.jan", "precip.apr", "precip.jul", "precip.oct", "log.con.percapita")

data_subset = data %>% 
  dplyr::select(all_of(variables))

#check

str(data_subset)

```


```{r correlation plot}
# variable correlation plot
library(ggcorrplot)

data2 = data %>% 
  dplyr::select("year", "spei", "pop", "dep.index", "built", "temp.jan", "temp.apr", "temp.jul", "temp.oct", "precip.jan", "precip.apr", "precip.jul", "precip.oct", "log.con.percapita", "total.con")

ggcorrplot(cor(data2),
           method = "square",
           type = "lower",
           lab = TRUE)

```
```{r}
library(janitor)

yr = tabyl(data, log.con.percapita, year)

```


```{r}

# train the random forest

set.seed(9371)

rf = randomForest(log.con.percapita ~ ., data = data_subset, ntree = 100)

print(rf)
```

# **Model Performance**

```{r}

rf$mse[rf$ntree] #mean of the squared residuals

rf.rmse = sqrt(rf$mse[rf$ntree]) # manually calculated RMSE

rf.rmse / diff(range(data$log.con.percapita)) #calculate the error based on the range of per capita conflict values

#~ 9% error when predicting the per capita conflict rate

```



# **iml predictor container**

```{r}

## Create a dataframe that has all of the features except the target

x = data_subset[which(names(data_subset) != "log.con.percapita")]

## Store the data and target in the Predictor() container
## parameters are the model - in this case rf
## the data - in this case the new data frame without target - x
## and the target 

predictor = Predictor$new(model = rf,
                          data = x,
                          y = data_subset$log.con.percapita)
```




```{r}
mp = ModelPerform

```



# **Feature Importance**


```{r, message=FALSE}

## Store the features in a FeatureImp object
## loss argument specifies the performance measure for error

imp = FeatureImp$new(predictor = predictor,
                     loss = "rmse")

```



```{r}

## Visualize the importance using ggplot2
library(ggplot2)

##Plot
plot(imp)


```


```{r}

# view importance results

imp$results

```


# **Feature Effects**



```{r, message=FALSE}

# Partial Dependence Plots
# population 

rf.pop = Partial$new(predictor = predictor,
                        feature = "pop",
                        aggregation = "pdp",
                        ice = TRUE)


rf.built = Partial$new(predictor = predictor,
                        feature = "built",
                        aggregation = "pdp",
                        ice = TRUE)


rf.dep = Partial$new(predictor = predictor,
                        feature = "dep.index",
                        aggregation = "pdp",
                        ice = TRUE)


rf.year = Partial$new(predictor = predictor,
                        feature = "year",
                        aggregation = "pdp",
                        ice = TRUE)

rf.aprtemp = Partial$new(predictor = predictor,
                        feature = "temp.apr",
                        aggregation = "pdp",
                        ice = TRUE)

```


```{r}

# center (this centers the impact of y hat on a starting value)
# in this case it centers it at the minimum value for and the curve is the deviation from this value
rf.pop$center(min(data_subset$pop))

# plot
p1 = plot(rf.pop) + ggtitle("Random Forest Population PDP")

p2 = plot(rf.built) + ggtitle("Random Forest Urbanization PDP")

p3 = plot(rf.dep) + ggtitle("Random Forest Deprivation PDP")

p4 = plot(rf.year) + ggtitle("Random Forest Year PDP")

p5 = plot(rf.aprtemp) + ggtitle("Random Forest Spring Temp PDP")



# center (this centers the impact of y hat on a starting value)
rf.pop$center(min(data$pop))
rf.dep$center(min(data$dep.index))
rf.aprtemp$center(min(data$temp.apr))
rf.built$center(min(data$built))
rf.year$center(min(data$year))

```


```{r message=FALSE}
library(gridExtra)
library(ggplot2)

# plot
p1 = plot(rf.pop) + ggtitle("Population PDP/ICE Plot")
p2 = plot(rf.built) + ggtitle("Urbanization PDP/ICE Plot")
p3 = plot(rf.dep) + ggtitle("Deprivation PDP/ICE Plot")
p4 = plot(rf.year) + ggtitle("Year PDP/ICE Plot")
p5 = plot(rf.aprtemp) + ggtitle("Spring Temperature PDP/ICE Plot")

gridExtra::grid.arrange(p1, p2, p3, p4, p5, nrow = 3)

```


# **Interactions**


```{r}
##Set up the interactions wrapper
##Play around with grid.size

interact = Interaction$new(predictor = predictor,
                           grid.size = 15)
```


```{r}

# visualize

plot(interact)
```


## **vivid Analysis**

```{r}
# visualization test
library(vivid)

set.seed(731)

viviRf2  <- vivi(fit = rf, 
                data = data_subset, #which data frame?
                response = "log.con.percapita",
                gridSize = 50,
                importanceType = "agnostic",
                nmax = 500,
                reorder = TRUE,
                predictFun = NULL,
                numPerm = 4,
                showVimpError = FALSE)


```


### **Heatmap Plot**

From Alan Inglis page - "The viviHeatmap function generates a heatmap that displays variable importance and interactions, with importance values on the diagonal and interaction values on the off-diagonal."


```{r}
# heatmap plot

viviHeatmap(mat = viviRf2,
            angle = 90, # rotate x-axis labels
            border = TRUE) # black border around the diagonal elements

```



### **Network Plot**

"With viviNetwork, a network graph is produced to visualize both importance and interactions."

```{r}
# network plot

# required packages
library(network) 
library(sna)
library(scales)
library(intergraph)

#visualize 
viviNetwork(mat = viviRf2)
```


```{r}
# remove nodes below a certain threshold

viviNetwork(mat = viviRf2,
            intThreshold = 0.05, # removes connections if the interaction threshold is below this value
            removeNode = TRUE) # removes nodes if the connection is below the threshold value
```

```{r}
library(lemon)

# extract the first five variables from our matrix to plot
#top5 <- colnames(viviRf)[1:5]
top4 = c("pop", "built", "year", "dep.index") 

pdpVars(data = data_subset,
        fit = rf,
        response = 'log.con.percapita',
        vars = top4,
        nIce = 1000) #number of ice curves plotted
```

```{r}

# extract the next variables to visualize

next2 = c("temp.apr", "temp.oct") 

pdpVars(data = data_subset,
        fit = rf,
        response = 'log.con.percapita',
        vars = next2,
        nIce = 1000) #number of ice curves plotted
```


**Generalized PDP Pairs**

"By employing a matrix layout, the pdpPairs function generates a generalized pairs partial dependence plot (GPDP) that encompasses univariate partial dependence (with ICE curves) on the diagonal, bivariate partial dependence on the upper diagonal, and a scatterplot of raw variable values on the lower diagonal, where all colours are assigned to points and ICE curves by the predicted ð‘¦Ì‚ value."



```{r}

#set.seed(1976)
#pdpPairs(data = pak2, 
         #fit =  ranger2.obj, 
         #response = "water", 
        # nmax = 500, 
         #gridSize = 10,         
         #vars = top4,
         #nIce = 100)

```








